Limb-darkening and exoplanets II: Choosing the Best Law for Minimum Bias Retrieval of Transit Parameters

### 2015-11-13 which_law_2_use.tex
revised version to be submitted

\title[Choosing the right law for minimum bias retrieval of transit parameters from transit lightcurves]
{Limb-darkening and exoplanets II: choosing the right law for minimum bias retrieval of transit parameters from transit lightcurves}
\author[Espinoza \& Jord\'an ]{
N\'estor Espinoza$^{1,2}$\thanks{E-mail:nespino@astro.puc.cl}, 
Andr\'es Jord\'an$^{2,1}$\thanks{E-mail: ajordan@astro.puc.cl}\\ 
$^{1}$Millennium Institute of Astrophysics, Vicu\~na Mackenna 4860, Santiago, Chile\\
$^{2}$Instituto de Astrof\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~na Mackenna 
4860, Santiago, Chile}
\begin{document}

@@

\begin{abstract}
It has recently been demonstrated that the parametrization of the limb-darkening effect on a star has a direct impact 
on the parameters retrieved from transit lightcurves such as the exoplanet's radius, scaled semi-major axis and inclination. However, 
studies regarding this issue have only focused on the widely used quadratic limb-darkening law, leaving outside of the analysis 
other proposed laws that in fact are better descriptors of model intensity profiles. In this work, we show that in fact 
laws such as the logarithmic, square-root and three-parameter do a better job at retrieving the mentioned parameters from transit 
lightcurves than the widely used quadratic law and, as such, we recommend using those instead. We also detail when to use each of those, 
which we note has a dependence on both stellar and transit parameters. In addition, we demonstrate that the exponential law should not be 
used as it fundamentally non-physical, typically producing a divergence towards negative intensities at the limb.
\end{abstract}

@@

\section{Introduction}
In the past decade, the study of transiting exoplanets has been evolving from discovery to very precise characterization of these systems, 
thanks to the exquisit precision allowed mainly by space-based observatories such as the Hubble Space Telescope (HST) and the Kepler mission. 
This in turn has allowed the exoplanet community to study precise mass-radius diagrams \citep[see, e.g.,][and references therein]{wm2014}, and in turn raise 
questions about the possibility of not only obtaining a precise determination of the internal composition of small, rocky exoplanets based on mass-radius
measurements \citep{dorn2015} or of their atmospheric fractions from radius measurements alone \citep{WL2015}, but also has allowed for the determination of
derived parameters of the systems that even allow for the obtention of stellar parameters directly from transit lightcurves through techniques such as asterodensity
profiling \citep{sm2003,kipping2014AP}, or of the detection of atmospheric features in exoplanet atmospheres through the technique of transmission spectroscopy. 
All these studies and techniques rely on the retrieval of precise and accurate transit parameters from transit lightcurves.

In a recent study \citep[][hereafter referred as EJ15]{ej2015}, we showed that the accuracy is actually catching up with the precision 
of these measurements due to our poor understanding of the limb-darkening effect. In particular, we showed that there are important 
biases both when fixing and fitting the limb-darkening coefficients in the light curve fitting process, with the latter arising from the fact 
that the popular and widely used quadratic law is unable to model the complex intensity profile of real stars and the former arising for this 
same reason plus the fact that different methods of fitting the model intensity profiles give rise to different limb-darkening coefficients and 
from the fact that we have an imperfect knowledge of the real intensity profiles of stars. This showed that fixing the limb-darkening coefficients is 
actually the worst option if one is willing to obtain precise \textit{and} accurate transit parameters, because the biases arise from three different sources, 
with the last one (the fact that we do not have a perfect modelling of the stellar intensity profiles of real stars) having an unknown but possibly 
large impact on them.

Given the above, unless the data quality is really poor, there is actually no good reason why one would actually want to fix the 
limb-darkening coefficients in the transit fitting process (assuming computational resources are not an issue). However, although fitting the 
limb-darkening coefficients seems to be a good solution to the accuracy problem, this still has the issue of the low flexibility of the quadratic limb-darkening 
law which still causes important biases on the retrieved transit parameters according to EJ15, which can be a large as $\sim 1\%$ for the planet-to-star 
radius ratio $R_p/R_*$, $\sim 2\%$ for the scaled semi-major axis, $a/R_*$, and $2\%$ for the inclination, where $R_p$ is the planetary radius, $R_*$ 
the stellar radius and $a$ the semi-major axis. The importance for missions like Kepler and for future missions like the Transiting Exoplanet Survey 
Satellite \cite[TESS,][]{ricker2014} is evident when looking at the most recent results from the Kepler mission: if we focus on the planet-to-star radius ratio 
alone, which according to EJ15 has an accuracy bias on the order of $\sim 0.2\%$, a query to the Nasa Exoplanet Archive\footnote{\url{http://exoplanetarchive.ipac.caltech.edu/}; query done on 29/09/2015.} shows that out of 3063 planetary candidates, 933 ($26\%$) have precisions better than 
$0.2\%$, and out of 1001 Kepler confirmed exoplanets, 463 (46\%) do too. This means that at least\footnote{It is important to note that this is an 
underestimate of the number of planet candidates with estimation errors, due to the fact that the Kepler pipeline makes use of fixed limb-darkening 
coefficients using the quadratic law \citep{rowe2015}, which as explained here and estimated in EJ15, gives rise to more severe biases than the one used for this 
estimation} $26\%$ of the planet candidates (from which, e.g., population studies, which rely on averaging out \textit{random} and not \textit{systematic} 
uncertainties like the ones introduced by limb-darkening) and almost half of the confirmed exoplanets (from which, e.g., characterization studies are 
based on) have important systematic errors.

An obvious solution to the above mentioned problem if one decides to fit the limb-darkening coefficients is to try to use \textit{other} laws to describe 
the intensity profile of stars. Although the non-linear law proposed by \cite{claret2000} seems to be the most flexible, the fact that it has four free parameters 
does not make it a very attractive choice in this scheme if one is willing to replace the quadratic law. However, laws with fewer parameters have been proposed, 
with the two-parameter laws being the exponential proposed by \cite{claret2003}, the logarithmic proposed by \cite{klinglesmith1970} and the square-root proposed 
by \cite{diazgimenez1992}. In addition, a very flexible three-parameter law was also proposed by \cite{sing2009}. These lower parameter laws seem to be 
very attractive for transit fitting purposes, due to their flexibility at following different intensity profiles \citep[see, e.g., ][for a comparison between the goodness of 
fit to ATLAS model atmospheres of the mentioned two-parameter laws which outperform the linear and quadratic laws in terms of following the intensity 
profiles]{howarth2011} and their low number of parameters.

Despite the attractive nature of the mentioned limb-darkening laws, testing their performance at retrieving transit parameters from transit 
lightcurves was a problem until recently for two reasons. First, there was no published algorithm that 
was capable of efficiently generating fast and accurate transit lightcurves using all of these non-standard laws. However, recently \cite{batman2015} published an 
algorithm that does exactly this called \texttt{batman}, enabling one then to generate transit lightcurves very efficiently with any arbitrary limb-darkening law. 
The second problem was that it was not clear how to sample limb-darkening coefficients in an informative (i.e., sampling all the physically possible parameter 
space) and efficient way for all of these laws. \cite{kipping2015} recently derived an algorithm to sample parameters from the three-parameter law by imposing physically plausible constrains on the intensity profiles which derived in constrains on the limb-darkening coefficients being fitted, while \cite{kipping2013}, using the same principle, derived an algorithmic way of 
doing this for two-parameter laws. However, he found that his algorithm was not applicable to the exponential law and thus no constrains for this law could 
be derived. In addition to this, the form of the logarithmic law used in \cite{kipping2013} differs from the standard one proposed by \cite{klinglesmith1970}, 
which is actually the one used in \texttt{batman}. Thus, in order to sample coefficients from this limb-darkening law, one first has to derive an efficient 
informative sampling scheme for this standard form of the law.

In this work, we aim at testing how well do these non-standard laws perform at retrieving transit parameters from transit lightcurves, and compare them to 
the retrievals done by the standard quadratic limb-darkening law. For this, we first follow \cite{kipping2013} in order to derive an efficient sampling scheme 
for the standard form of the logarithmic and exponential laws, and show that the latter is actually \textit{always} non-physical in the sense proposed by 
\cite{kipping2013}. Then, we simulate transit lightcurves and perform a simulation study similar to the one done in EJ15 using the derived sampling schemes 
for the logarithmic law and the ones detailed in \cite{kipping2013} for the other two-parameter laws and \cite{kipping2015} for the three-parameter law, 
detailing when one law should be preferred over the other.

This work is organised as follows. In \S2, we first revisit the logarithmic and exponential 
laws in order to try to derive an efficient informative sampling strategy following the methods in \cite{kipping2013} for the standard forms of these laws. 
In \S3 we use those results in addition to the ones published by \cite{kipping2013} and \cite{kipping2015} in order to simulate transit lightcurves to test 
how well these laws perform in retrieving transit parameters. In \S4 we present a discussion of our work and present our conclusions in \S5.

@@

As explained in the introduction, an informative and efficient sampling of limb-darkening coefficients 
in order to fit them to transit lightcurves is fundamental. Motivated by this, \cite{kipping2013} introduced 
an algorithm which first imposes physically plausible constrains on the intensity profiles, namely, an everywhere 
positive intensity profile ($I(\mu)>0$) and a decreasing intensity profile from center to limb ($\partial I(\mu)/\partial \mu>0$), 
which in turn impose constrains on the parameters of the studied two-parameter law. With these constrains at hand, 
\cite{kipping2013} devised a triangular sampling strategy which only require sampling two uniformly distributed numbers between 
0 and 1, $q_1$ and $q_2$, which through a transformation can be converted in order to sample coefficients that follow the 
derived constrains. This was done in that work for the most popular two-parameter limb-darkening laws with the exception of the 
exponential limb-darkening law, whose derived constrains on the coefficients were not enough 
to use the triangular sampling technique, and for the typical form of the logarithmic law, which is 
different to the one used in \cite{kipping2013}.

We here first derive an informative and efficient sampling strategy for the typical form of the logarithmic law and then 
we study the exponential law, for which we conclude that deriving an efficient sampling strategy is actually 
\textit{impossible}; this implies that the law is fundamentally non-physical and, therefore, should not be used.

\subsection{Efficient sampling from the logarithmic law}
As described in \cite{klinglesmith1970}, the logarithmic law is given by
\begin{eqnarray}
\label{loglaw} I(\mu) = 1 - l_1(1-\mu)-l_2\mu \ln\mu.
\end{eqnarray}
However, \cite{kipping2013} derived the constrains for a law similar (but 
not the same) to the original logarithmic law, given by
\begin{eqnarray*}
I_K(\mu) = 1 - A(1-\mu)-B\mu (1-\ln\mu),
\end{eqnarray*}
which can be viewed as the original logarithmic law with an extra linear term. Given that this 
is not the original law many authors have adopted both to obtain limb-darkening coefficients 
and to model transit lightcurves, we here derive a sampling strategy for the original logarithmic 
limb-darkening law given in eq. (\ref{loglaw}). To do this, we start by imposing 
physically plausible constrains on the intensity profiles. The constrain of an everywhere positive intensity profile implies
\begin{eqnarray*}
 l_1(1-\mu)+l_2\mu \ln\mu < 1  &\forall& 0<\mu<1.
\end{eqnarray*}
In order to ensure this condition, we need to find the extrema of the expression in the 
left-hand side. We note that the maximum of the expression is obtained only if $l_2<0$, 
while the minimum of the expression is obtained if $l_2>0$. In both cases, this happens 
at $\mu_e=\exp [(l_1-l_2)/l_2]$, where we have assumed $l_2 \neq 0$. If $l_2=0$, then 
we obtain the inequality $l_1(1-\mu)<1$, which implies $l_1<1$, constraint we will see 
is valid for all the possible values of $l_2$. In the case $l_2>0$, because 
$\mu_e$ gives a minimum, 
$\mu\to 1$ and $\mu \to 0+$ maximise the expression in the desired range. The first limit 
gives the trivial constrain $1>0$, while the second limit gives the constrain $l_1<1$. In 
the case $l_2<0$, replacing $\mu_e$ in the expression leads to the constrain
\begin{eqnarray*}
\frac{l_1-1}{\mu_e} < l_2,
\end{eqnarray*}  
which, because $l_2<0$, again implies $l_1<1$. The condition 
for a decreasing intensity profile from center to limb, on the other hand, leads to the 
condition
\begin{eqnarray*}
l_1-l_2(1+\ln\mu) > 0  &\forall& 0<\mu<1.
\end{eqnarray*}
The left-hand side has, again, a division depending on the sign of $l_2$. If $l_2>0$, then 
the expression is convex and has no absolute minimum. It can be seen that then the 
expression is always satisfied for $0<\mu < 1$. In the limit $\mu \to 1$, the expression leads 
to the condition $l_1>l_2$. For the $\mu \to 0^+$ limit, however, no condition can be derived 
because the expression diverges. If $l_2<0$, on the other hand, the expression is concave, and 
it does not have an absolute maximum. It can be seen, however, that the expression in the limit 
$\mu \to 1$, leads to $l_1>l_2$ just as in the first case, but in this case the expression is not valid 
for $\mu \to 0^+$, because it goes to $-\infty$. This implies that for $l_2<0$ the inequality will 
never be satisfied in the desired range and, thus, we need $l_2>0$ in order for the profile to be 
everywhere decreasing. In summary, the derived condition for an everywhere positive intensity profile is
\begin{eqnarray*}
l_1<1,
\end{eqnarray*}
while the conditions for an everywhere decreasing intensity profile from center to limb are
\begin{eqnarray*}
l_2>0,\\
l_1>l_2.
\end{eqnarray*}

@@

In order to use the triangular sampling technique, one needs to re-parametrize 
the constrains in order for them to be sampled on a right-angled triangle with this angle posed in the origin. A 
parametrisation that does this is the one with
\begin{eqnarray}
\label{v1}&v_1& = 1-l_1,\\
\label{v2}&v_2& = l_2.
\end{eqnarray}
If we now consider the transformations \citep[see][]{kipping2013}
\begin{eqnarray}
\label{v1q} &v_1& = \sqrt{q_1}q_2,\\
\label{v2q} &v_2& = 1 - \sqrt{q_1},
\end{eqnarray}
sampling $q_1$ and $q_2$ from uniform distributions between $(0,1)$ leads to a sampling with the desired constrains 
between $l_1$ and $l_2$. Replacing the expressions in eqs. (\ref{v1q}) and (\ref{v2q}) in eqs. (\ref{v1}) and (\ref{v2}) gives 
\begin{eqnarray*}
l_1 &=& 1-\sqrt{q_1}q_2\\
l_2 &=& 1-\sqrt{q_1}
\end{eqnarray*}
which leads to the inverse equations
\begin{eqnarray*}
q_1 &=& (1-l_2)^2,\\
q_2 &=& \frac{1-l_1}{1-l_2}.
\end{eqnarray*}
We note that, as expected, these relations return limb-darkening coefficients that are physically plausible for this law (i.e., they follow 
the derived constrains). Furthermore, we note these relations are different to the ones derived in \cite{kipping2013}; these are only applicable 
for his form of the logarithmic law.

\subsection{The exponential law is non-physical}
We now study the exponential limb-darkening law. This law was introduced by \cite{claret2003}, and is given by
\begin{eqnarray*}
I(\mu) &=& 1-e_1(1-\mu)-e_2/(1-e^\mu).
\end{eqnarray*}
\cite{kipping2013} tried to apply the same methods used for the other two-parameter laws in order to 
derive constrains and hence an efficient sampling scheme for the coefficients of this law, but observed that the triangular 
sampling technique was not applicable in this case because the imposed physical constrains were not 
able to yield a sufficient number of relations between the coefficients for it to 
be used. However, apparently overlooked by \cite{kipping2013} is the fact that the exponential law will \textit{never} 
yield physically plausible coefficients for $0<\mu<1$, because the conditions of an everywhere positive and decreasing 
intensity profile from center to limb cannot be satisfied at the same time, as we now show.

We start by imposing an everywhere positive intensity profile for this law. This leads to the relation
\begin{eqnarray*}
e_1(1-\mu)+e_2/(1-e^\mu)<1 &\forall& 0<\mu<1,
\end{eqnarray*}
and the objective is to study the left-hand side. We note that this expression has a minimum if 
$e_2<0$, while it has a maximum when $e_2>0$. We see that in the first case, as $\mu \to 0^+$, 
the expression tends to $\infty$, and therefore the expression is never satisfied; this implies that 
$e_2$ cannot be less than zero in order to have an everywhere positive intensity profile. On the 
other hand, if we impose an everywhere decreasing intensity profile for this law from center to 
limb, one can note this leads to the constrain:
\begin{eqnarray*}
e_1 - e_2 \frac{e^\mu}{(1-e^\mu)^2}>0 &\forall& 0<\mu<1.
\end{eqnarray*}
The form of the left-hand side expression again depends on the value of $e_2$ in the same way 
as before; however, in this case there is no absolute maximum or minimum and, thus, it suffices 
to study the expression at the borders. We note that in order for the expression to be satisfied 
for all the values of $\mu$, $e_2<0$, because only in this case the expression goes to $\infty$ 
when $\mu \to 0^+$. Because in order for the exponential law to have an everywhere positive intensity 
profile this condition was ruled out, this implies that both conditions cannot be met at the 
same time and, thus, there is no combination of coefficients $(e_1,e_2)$ that satisfy the physically 
plausible conditions suggested by \cite{kipping2013} for this law: in this sense, the exponential 
law introduced by \cite{claret2003} is non-physical.

@@

\section{Biases in transit parameters using two-parameter laws}
Having solved the issues regarding the sampling of limb-darkening coefficients from the 
logarithmic and exponential laws, we now turn to the study of the 
biases introduced on retrieved transit parameters in the transit fitting process introduced by using different limb-darkening laws. In 
order to perform this study, we perform a similar study as the one done in EJ15, but now 
(1) using the \texttt{batman} transit code described in \cite{batman2015}, (2) using three different two-parameter 
laws: the quadratic, logarithmic and square-root laws (with the exponential law left out of our analysis because, 
as was shown in \S 2.2, it's fundamentally non-physical) and the three-parameter law and (3) smaller steps in 
the planet-to-star radius ratio, $p=R_p/R_*$, in order to illustrate the evolution of the biases with this parameter 
in a cleaner way: small ratios ($p=0.01$), medium ratios ($p=0.07$) and large ones ($p=0.13$) .

>>>>

\title[Choosing the best law for minimum bias retrieval of transit parameters]
{Limb-darkening and exoplanets II: Choosing the Best Law for Minimum Bias Retrieval of Transit Parameters}
\author[Espinoza \& Jord\'an ]{
N\'estor Espinoza$^{1,2}$\thanks{E-mail:nespino@astro.puc.cl}, 
Andr\'es Jord\'an$^{1,2}$\thanks{E-mail:ajordan@astro.puc.cl}\\ 
$^{1}$Instituto de Astrof\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~na Mackenna 
4860, Santiago, Chile
$^{2}$Millennium Institute of Astrophysics, Santiago, Chile\\
}
\begin{document}

@@

\begin{abstract}
The parametrization of the limb-darkening effect of a star has a direct impact 
on the parameters retrieved from transit lightcurves such as the exoplanet's radius, scaled semi-major axis and inclination. Studies of the effect of limb darkening assumptions on the retrieved parameters have mostly focused on the widely used quadratic limb-darkening law, leaving out 
other proposed laws that are in fact better descriptions of model intensity profiles. In this work, we show that 
laws such as the logarithmic, square-root and three-parameter laws do a better job at retrieving the mentioned parameters from transit 
lightcurves than the widely used quadratic law and therefore we recommend using those instead. We also provide code to guide the decision of when to use each of the laws,  which we note has a dependence on both stellar and transit parameters. Finally, we demonstrate that the so-called exponential law is non-physical as it typically produces negative intensities close to the limb and should therefore not be used.
\end{abstract}

@@

\section{Introduction}

In the past decade, the study of transiting exoplanets has evolved from discovery to very precise characterization of these systems, 
thanks to the exquisite precision allowed mainly by space-based observatories such as the {\em Hubble Space Telescope} ({\em HST}) and the {\em Kepler} mission \citep{koch2010}.
This has allowed the study of precise mass-radius diagrams \citep[see, e.g.,][and references therein]{wm2014}, and in turn  allow the possibility of not only obtaining a precise determination of the internal composition of small, rocky exoplanets based on mass-radius
measurements \citep{dorn2015} or of their atmospheric fractions from radius measurements alone \citep{WL2015}, but also of the determination of
derived parameters of the systems that even allow for the estimation of stellar parameters directly from transit lightcurves through techniques such as asterodensity
profiling \citep{sm2003,kipping2014AP}, or of the detection of atmospheric features in exoplanet atmospheres through the technique of transmission spectroscopy. 
All these studies and techniques rely on the retrieval of precise and accurate transit parameters from transit lightcurves.

In a recent study \citep[][hereon EJ15]{ej2015}, we showed that the accuracy is actually catching up with the precision 
of these measurements due to our imperfect understanding of the limb-darkening effect. In particular, we showed that there are important 
biases both when fixing and fitting for the limb-darkening coefficients in the light curve fitting process. The biases when fitting for the limb darkening coefficients arise from the fact 
that the popular and widely used quadratic law is unable to model the complex intensity profile of real stars,  and the biases when fixing the coefficients arise for the
same reason plus the facts that different methods of fitting the model intensity profiles (transit light curve versus intensity profile fitting) give rise to different limb-darkening coefficients and 
the fact that we have an imperfect knowledge of the real intensity profiles of stars. We showed that fixing the limb-darkening coefficients is 
the worst option if one aims to obtain precise \textit{and} accurate transit parameters, because the biases arise from three different sources, 
with the last one (the fact that we do not have a perfect modelling of the stellar intensity profiles of real stars) having an unknown but possibly 
large impact.

Given the above, unless the data quality is really poor (and one is willing to trade bias for variance), there is actually no good reason to fix the limb-darkening coefficients in the transit fitting process (assuming computational resources are not an issue). However, although fitting the 
limb-darkening coefficients seems to be a good solution to the accuracy problem, there remains the issue of the low flexibility of the quadratic limb-darkening 
law which can cause important biases on the retrieved transit parameters as shown in EJ15. These biases can be a large as $\sim 1\%$ for the planet-to-star 
radius ratio $R_p/R_*$, $\sim 2\%$ for the scaled semi-major axis, $a/R_*$, and $2\%$ for the inclination, where $R_p$ is the planetary radius, $R_*$ 
the stellar radius and $a$ the semi-major axis. The importance for missions like {\em Kepler} and for future missions like the Transiting Exoplanet Survey 
Satellite \cite[TESS,][]{ricker2014} is evident when looking at the most recent results from the {\em Kepler} mission: if we focus on the planet-to-star radius ratio measurements 
alone, which according to EJ15 have an accuracy bias on the order of $\sim 0.2\%$, a query to the Nasa Exoplanet Archive\footnote{\url{http://exoplanetarchive.ipac.caltech.edu/}; query done on 29/09/2015.} shows that out of 3063 planetary candidates, 933 ($26\%$) have precisions better than 
$0.2\%$, and out of 1001 {\em Kepler} confirmed exoplanets, 463 (46\%) do too. This means that at least\footnote{It is important to note that this is an 
underestimate of the number of planet candidates with estimation errors, due to the fact that the Kepler pipeline makes use of fixed limb-darkening 
coefficients using the quadratic law \citep{rowe2015}, which as explained here and estimated in EJ15, gives rise to more severe biases than the one used for this 
estimation} $26\%$ of the planet candidates (from which, e.g., population studies, which rely on averaging out \textit{random} and not \textit{systematic} 
uncertainties like the ones introduced by limb-darkening) and almost half of the confirmed exoplanets (from which, e.g., characterization studies are 
based on) have significant systematic errors.

An obvious solution to the issues mentioned above if one decides to fit the limb-darkening coefficients is to try to use alternative laws to describe 
the intensity profile of stars. Although the non-linear law proposed by \cite{claret2000} seems to be the most flexible, the fact that it has four free parameters 
does not make it a very attractive. Several laws with fewer parameters have been proposed. Two-parameter laws are the exponential proposed by \cite{claret2003}, the logarithmic proposed by \cite{klinglesmith1970} and the square-root proposed 
by \cite{diazgimenez1992}. In addition, a very flexible three-parameter law was also proposed by \cite{sing2009}. These laws with fewer parameters are potentially 
very attractive for transit fitting purposes, due to their flexibility at following different intensity profiles  and their low number of parameters \citep[see, e.g., ][for a comparison between the goodness of 
fit to ATLAS model atmospheres of the mentioned two-parameter laws which outperform the linear and quadratic laws in terms of following the intensity 
profiles]{howarth2011}.

Despite the attractive nature of the above mentioned limb-darkening laws, testing their performance at retrieving transit parameters from transit 
lightcurves was a problem until recently for two reasons. First, there was no publicly available code capable of efficiently generating fast and accurate transit lightcurves using all of these non-standard laws. However, recently \cite{batman2015} published an 
algorithm that does exactly this called \texttt{batman}, enabling the generation of transit lightcurves very efficiently with any arbitrary limb-darkening law. 
The second problem was that it was not clear how to sample limb-darkening coefficients in an informative (i.e., sampling all the physically possible parameter 
space) and efficient way for all of these laws. Recently, \cite{kipping2015} presented an algorithm to sample parameters from the three-parameter law by imposing physically motivated constraints on the intensity profiles which derived in constrains on the limb-darkening coefficients being fitted. cite{kipping2013}, using the same principles, derived an algorithmic way of 
doing this for two-parameter laws. He found that his algorithm was not applicable to the exponential law and thus no constrains for this law could 
be derived. We note that the form of the logarithmic law used in \cite{kipping2013} differs from the standard one proposed by 
\cite{klinglesmith1970}. Thus, in order to  efficiently sample coefficients from the latter limb-darkening law one has to derive an efficient  informative sampling scheme for it.

In this work, we extend the work of EJ15 by testing how well these non-standard laws perform at retrieving transit parameters from transit lightcurves, and compare them to 
the retrievals done by the standard quadratic limb-darkening law. We follow \cite{kipping2013} in order to derive an efficient sampling scheme 
for the standard form of the logarithmic and exponential laws, and show that the latter is actually \textit{always} non-physical in the sense proposed by 
\cite{kipping2013}. Then, we simulate transit lightcurves and perform a simulation study similar to the one done in EJ15 using the derived sampling schemes 
for the logarithmic law and the sampling schemes detailed in \cite{kipping2013} for the other two-parameter laws and \cite{kipping2015} for the three-parameter law. We use our simulations to provide guidance as to when a particular law should be preferred.

This work is organised as follows. In \S2, we revisit the logarithmic and exponential 
laws in order to derive an efficient informative sampling strategy following the methods in \cite{kipping2013} for the standard forms of these laws. 
In \S3 we use those results in addition to the ones published by \cite{kipping2013} and \cite{kipping2015} in order to simulate transit lightcurves to test 
how well these laws perform in retrieving transit parameters. In \S4 we present a discussion and conclusions of our work.

@@

As explained in the introduction, an informative and efficient sampling of limb-darkening coefficients 
in order to fit them to transit lightcurves is fundamental. Motivated by this, \cite{kipping2013} introduced 
a procedure which first imposes physically motivated constrains on the intensity profiles, namely, an everywhere 
positive intensity, $I(\mu)>0$,  and a decreasing intensity profile from center to limb, $\partial I(\mu)/\partial \mu>0$, 
which in turn impose constraints on the two parameters of the law. With these constraints in hand, 
\cite{kipping2013} devised a triangular sampling strategy which only requires sampling two uniformly distributed numbers between 
0 and 1, $q_1$ and $q_2$, which through a transformation can be transformed to sample coefficients. This was done in that work for the most popular two-parameter limb-darkening laws with the exceptions of the 
exponential limb-darkening law, whose derived constraints on the coefficients were not enough 
to use the triangular sampling technique, and the typical form of the logarithmic law, which is 
different to the one used in \cite{kipping2013}.

We now derive an informative and efficient sampling strategy for the more typical form of the logarithmic law and then 
we study the exponential law, for which we conclude that deriving an efficient sampling strategy is actually 
impossible; this implies that the law is fundamentally non-physical and, therefore, should not be used.

\subsection{Efficient sampling from the logarithmic law}

As described in \cite{klinglesmith1970}, the logarithmic law is given by
\begin{eqnarray}
\label{loglaw} I(\mu) = 1 - l_1(1-\mu)-l_2\mu \ln\mu.
\end{eqnarray}
However, \cite{kipping2013} derived the constrains for a law similar (but 
not the same) to the original logarithmic law, given by
\begin{eqnarray*}
I_K(\mu) = 1 - A(1-\mu)-B\mu (1-\ln\mu).
\end{eqnarray*}

Given that this 
is not the law many authors have adopted we  derive here a sampling strategy for the original law given in eq. (\ref{loglaw}).  The constrain of an everywhere positive intensity profile implies
\begin{eqnarray*}
 l_1(1-\mu)+l_2\mu \ln\mu < 1  &\forall& 0<\mu<1.
\end{eqnarray*}
In order to ensure this condition, we need to find the extrema of the expression on the 
left-hand side. We note that a maximum of the expression is obtained only if $l_2<0$, 
while a minimum of the expression is obtained if $l_2>0$. In both cases, this happens 
at $\mu_e=\exp [(l_1-l_2)/l_2]$, where we have assumed $l_2 \neq 0$. If $l_2=0$, then 
we obtain the inequality $l_1(1-\mu)<1$, which implies $l_1<1$, a constraint we will see 
is valid for all the possible values of $l_2$. In the case $l_2>0$, because 
$\mu_e$ gives a minimum, 
$\mu\to 1$ and $\mu \to 0+$ maximise the expression in the desired range. The first limit 
gives the trivial constrain $1>0$, while the second limit gives the constrain $l_1<1$. In 
the case $l_2<0$, replacing $\mu_e$ in the expression leads to the constrain
\begin{eqnarray*}
\frac{l_1-1}{\mu_e} < l_2,
\end{eqnarray*}  
which, because $l_2<0$, again implies $l_1<1$. The condition 
for a decreasing intensity profile from center to limb, on the other hand, leads to the 
condition
\begin{eqnarray*}
l_1-l_2(1+\ln\mu) > 0  &\forall& 0<\mu<1.
\end{eqnarray*}
The left-hand side has different behaviour depending on the sign of $l_2$. If $l_2>0$, then 
the expression is convex and has no absolute minimum. It can be seen then that  the 
expression is always satisfied for $0<\mu < 1$. In the limit $\mu \to 1$, the expression leads 
to the condition $l_1>l_2$. For the $\mu \to 0^+$ limit no condition can be derived 
because the expression diverges. If $l_2<0$ the expression is concave, and 
it does not have an absolute maximum in $]0,1[$, the expression in the limit 
$\mu \to 1$ leads to $l_1>l_2$ just as in the first case, but in this case the expression is not valid 
for $\mu \to 0^+$, because it goes to $-\infty$. This implies that for $l_2<0$ the inequality will 
never be satisfied in the desired range and, thus, we need $l_2>0$ in order for the profile to be 
everywhere decreasing. In summary, the derived condition for an everywhere positive intensity profile is
\begin{eqnarray*}
l_1<1,
\end{eqnarray*}
while the conditions for an everywhere decreasing intensity profile from center to limb are
\begin{eqnarray*}
l_2>0,\\
l_1>l_2.
\end{eqnarray*}

@@

In order to use the triangular sampling technique, one needs to re-parametrize 
the constrains in order for them to be sampled on a right-angled triangle with this angle posed at the origin. A 
parametrisation that does this is the one with
\begin{eqnarray}
\label{v1}&v_1& = 1-l_1,\\
\label{v2}&v_2& = l_2.
\end{eqnarray}
If we now consider the transformations \citep[see][]{kipping2013}
\begin{eqnarray}
\label{v1q} &v_1& = \sqrt{q_1}q_2,\\
\label{v2q} &v_2& = 1 - \sqrt{q_1},
\end{eqnarray}
sampling $q_1$ and $q_2$ from uniform distributions between $(0,1)$ leads to a sampling with the desired constrains. Replacing the expressions in eqs. (\ref{v1q}) and (\ref{v2q}) in eqs. (\ref{v1}) and (\ref{v2}) gives 
\begin{eqnarray*}
l_1 &=& 1-\sqrt{q_1}q_2\\
l_2 &=& 1-\sqrt{q_1}
\end{eqnarray*}
which leads to the inverse equations
\begin{eqnarray*}
q_1 &=& (1-l_2)^2,\\
q_2 &=& \frac{1-l_1}{1-l_2}.
\end{eqnarray*}
We note that, as expected by construction, these relations return limb-darkening coefficients that are physically plausible for this law. Furthermore, we note these relations are different to the ones derived 
in \cite{kipping2013} as those are only applicable 
for his adopted form of the logarithmic law.

\subsection{The exponential law is non-physical}

We now study the exponential limb-darkening law. This law was 
introduced by \cite{claret2003}, and is given by
\begin{eqnarray*}
I(\mu) &=& 1-e_1(1-\mu)-e_2/(1-e^\mu).
\end{eqnarray*}
\cite{kipping2013} tried to apply the same methods used for the other two-parameter laws but observed that the triangular 
sampling technique was not applicable in this case because the physical constraints imposed were not 
able to yield a sufficient number of relations between the coefficients. However, apparently overlooked by \cite{kipping2013} is the fact that the exponential law will never
yield physically plausible coefficients for $0<\mu<1$, because the conditions of an everywhere positive and decreasing 
intensity profile from center to limb cannot be satisfied at the same time, as we now show.

We start by imposing an everywhere positive intensity profile for this law, which leads to the relation
\begin{eqnarray*}
e_1(1-\mu)+e_2/(1-e^\mu)<1 &\forall& 0<\mu<1,
\end{eqnarray*}
and the objective is to study the left-hand side. We note that this expression has a minimum if 
$e_2<0$, while it has a maximum if $e_2>0$. We see that in the first case, 
as $\mu \to 0^+$, 
the expression tends to $+\infty$, and therefore the expression cannot be satisfied for all $\mu$; this implies that 
$e_2$ cannot be less than zero. On the 
other hand, if we impose an everywhere decreasing intensity profile for this law from center to 
limb, this leads to the constraint
\begin{eqnarray*}
e_1 - e_2 \frac{e^\mu}{(1-e^\mu)^2}>0 &\forall& 0<\mu<1.
\end{eqnarray*}
The form of the left-hand side expression again depends on the value of $e_2$ in the same way 
as before, however, in this case there is no absolute maximum or minimum in $]0,1[$] and, thus, it suffices 
to study the expression at the borders. We note that in order for the expression to be satisfied 
for all the values of $\mu$, $e_2<0$, because only in this case the expression goes to $\infty$ 
when $\mu \to 0^+$. Because in order for the exponential law to have an everywhere positive intensity 
profile this condition was ruled out, this implies that both conditions cannot be met at the 
same time and, thus, there is no combination of coefficients $(e_1,e_2)$ that satisfy the physically 
plausible conditions suggested by \cite{kipping2013} for this law. This implies that the exponential 
law introduced by \cite{claret2003} is non-physical.

@@

\section{Biases in transit parameters using two-parameter laws}

Having dealt with the sampling of limb-darkening coefficients from the 
logarithmic and exponential laws, we now turn to the study of the 
biases introduced on retrieved transit parameters in the transit fitting process introduced by using different limb-darkening laws. We perform a similar study as the one done in EJ15, but now 
(1) using the \texttt{batman} transit code described in \cite{batman2015}, (2) using three different two-parameter 
laws: the quadratic, logarithmic and square-root laws, and the three-parameter law and (3) smaller steps in 
the planet-to-star radius ratio, $p=R_p/R_*$, in order to illustrate the evolution of the biases with this parameter 
in a cleaner way: small ratios ($p=0.01$), medium ratios ($p=0.07$) and large ones ($p=0.13$).
